---
title: "Group Project 4 COVID"
author: 'covid_p11'
date: '2022-03-28'
output:
  html_document:
    toc: true
    number_sections: true
    theme: flatly
    highlight: tango
    toc_float: true
    toc_depth: 4
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(maps)
library(plotly)
library(caret)
library(dplyr)
library(lubridate)
library(scales)
library(magrittr)
library(tidyr)
library(pkr)
library(plotly)
library(ggthemes)
library(tseries)
library(quantmod)
library(roll)
library(plotly)
library(gganimate)
library(tidyquant)
library(caret)
library(klaR)
```

Topic: COVID-19 New cases
visualisation dashboard: Shiny app

# NEW FINAL JOIN CODE BY MAXIM

```{r}
##########################
# THIS IS THE FINAL JOIN #
#######################################################################
# use covid_joined for your work and filter it to country if you need #
#######################################################################

covid = read.csv("owid-covid-data.csv")
sp500 = read.csv("SPY Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for US
TOPIX = read.csv("TOPIX Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for Japan
ASX200 = read.csv("S&P_ASX 200 Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for Australia
NSEI = read.csv("Nifty 50 Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for India
SSEC = read.csv("Shanghai Composite Historical Data.csv", fileEncoding = 'UTF-8-BOM') # this is for China

covid2 <- covid

sp <- sp500
sp$Change.. <- as.numeric(gsub("%","",sp$Change..))
max = max(sp$Change..)
max

covid2$date = as.character(ymd(covid2$date))
sp500$Date <- as.character(mdy(sp500$Date))
TOPIX$Date <- as.character(mdy(TOPIX$Date))
ASX200$Date <- as.character(mdy(ASX200$Date))
NSEI$Date <- as.character(mdy(NSEI$Date))
SSEC$Date <- as.character(mdy(SSEC$Date))

colnames(covid2)[4] = "Date"

covid_US <- covid2 %>% filter(location == "United States")
covid_AUS <- covid2 %>% filter(location == "Australia")
covid_IND <- covid2 %>% filter(location == "India")
covid_JPN <- covid2 %>% filter(location == "Japan")
covid_CHN <- covid2 %>% filter(location == "China")

covid_US

df_1 = inner_join(sp500, covid_US, by = "Date")
df_2 = inner_join(TOPIX, covid_JPN, by = "Date")
df_3 = inner_join(ASX200, covid_AUS, by = "Date")
df_4 = inner_join(NSEI, covid_IND, by = "Date")
df_5 = inner_join(SSEC, covid_CHN, by = "Date")

covid_joined <- rbind(df_1, df_2, df_3, df_4, df_5)

covid_joined$Date <- ymd(covid_joined$Date)
covid_joined$Price <- as.numeric(gsub(",","",covid_joined$Price))
covid_joined$Open <- as.numeric(gsub(",","",covid_joined$Open))
covid_joined$High <- as.numeric(gsub(",","",covid_joined$High))
covid_joined$Low <- as.numeric(gsub(",","",covid_joined$Low))
covid_joined$Change.. <- as.numeric(gsub("%","",covid_joined$Change..))

# example analysis for US
US_data = covid_joined[covid_joined$location == "United States",]
plot(US_data$Price ~ US_data$Date)
```

# Question 1 - intial data visualisation

## First visualisation 

- visualising distribution of gains and losses amoung different years

```{r}
covid_joined <- covid_joined %>% 
  filter(Date >= "2020-03-11") %>%
  mutate(Year_category = case_when( 
    Date >= "2020-03-11" & Date < "2021-01-01" ~ "1st year", 
    Date >= "2021-01-01" & Date < "2022-01-01" ~ "2nd year", 
    Date >= "2022-01-01" & Date < "2023-01-01" ~ "3rd year (ongoing)"))

# not using the ggplot but leaving it here in case, instead using the plotly becuase it is interactive

#ggplot(covid_joined, aes(x=Change.., y = location, fill = Year_category)) + geom_boxplot(outlier.colour="black",
#             outlier.size=0.3, notch=TRUE) + theme_clean() + labs(y = "country", x = "Distribution of daily #gains/losses (%)") + 
#ggtitle("Distrubtion of daily gains/losses during the years of the pandemic") 

fig <- plot_ly(covid_joined, x = ~Change.., y = ~location, color = ~Year_category, type = "box") %>% layout(boxmode = "group", title = 'Distribution of gains and losses over the pandemic',
         xaxis = list(showgrid = TRUE),
         yaxis = list(showgrid = TRUE))
unique(covid_joined$location)
```

## evolution of countries through the pandemic (not finisihed)

```{r, eval=FALSE}
head(covid_joined)

# Asset tickers

tickers = c("SPY")

price_data <- tq_get(tickers,
                     from = '2017-12-27',
                     to = '2021-12-27',
                     get = 'stock.prices')

price_data

ret_data <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "ret")


port_ret_tidyquant <- ret_data %>%
  tq_portfolio(assets_col = symbol,
               returns_col = ret,
               col_rename = 'port_ret',
               geometric = FALSE)



covid_joined$Change.. <- as.numeric(gsub("%","",covid_joined$Change..))



sp500$cr <- (as.numeric(gsub("%","",sp500$Change..))/100) %>% mutate(cr = cumprod(1 + sp500$cr))

port_cumulative_ret_tidyquant

plot(port_cumulative_ret_tidyquant$port_ret~port_cumulative_ret_tidyquant$date)

port_ret_tidyquant <- ret_data %>%
  tq_portfolio(assets_col = symbol,
               returns_col = ret,
               weights = wts,
               col_rename = 'port_ret',
               geometric = FALSE)

port_cumulative_ret_tidyquant <- port_ret_tidyquant %>%
  mutate(cr = cumprod(1 + port_ret))
```

## not finished

```{r, eval=FALSE}
library(tidyquant)

# Asset tickers

tickers = c('BND', 'VB', 'VEA', 'VOO', 'VWO')

# Asset weights

wts = c(0.1,0.2,0.25,0.25,0.2)

price_data <- tq_get(tickers,
                     from = '2013-01-01',
                     to = '2018-03-01',
                     get = 'stock.prices')

ret_data <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "ret")

port_ret_tidyquant <- ret_data %>%
  tq_portfolio(assets_col = symbol,
               returns_col = ret,
               weights = wts,
               col_rename = 'port_ret',
               geometric = FALSE)

ret_data <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "ret")

wts_tbl <- tibble(symbol = tickers,
                  wts = wts)
ret_data <- left_join(ret_data,wts_tbl, by = 'symbol')

ret_data <- ret_data %>%
  mutate(wt_return = wts * ret)

port_ret <- ret_data %>%
  group_by(date) %>%
  summarise(port_ret = sum(wt_return))

port_cumulative_ret <- port_ret %>%
  mutate(cr = cumprod(1 + port_ret))

port_cumulative_ret_tidyquant <- port_ret %>%
  mutate(cr = cumprod(1 + port_ret))


port_ret_tidyquant
port_ret

```


```{r, eval=FALSE}
covid_joined_vis %>% 
  ggplot(aes(total_cases, total_stock, size=pop, colour = country)) +
  geom_point(alpha = 0.5, show.legend = TRUE) +
  scale_size(range = c(2,12)) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita in USD', y = 'Life Expectancy') +
  transition_time(year) +
  shadow_wake(.3)+
  ease_aes('linear')
```


## CLEANING AND (OLD JOIN CODE) 

- (not going to use but will leave it because i think some of the code currently depends on it, will fix in the meeting together)

```{r}
covid = read.csv("owid-covid-data.csv")
ASX200 = read.csv("S&P_ASX 200 Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for Australia
```

# Q2

```{r}
df_q2 = covid_joined

colnames(df_q2)[7] = 'Change'

write.csv(df_q2,"df_q2.csv")
```

## Rolling correlation 

- (part of question 2 done by maxim) BTW THIS CODE IS REALLY MESSY :)

```{r, eval=FALSE}
mySymbols <- c('AMZN', '^GSPC','NFLX','MSFT','GOOG','NVAX','DIS','WMT','COST','V','FB','AXP','CMCSA','PFE','MRNA')
myStocks <-lapply(mySymbols, function(x) {getSymbols(x, src = "yahoo", 
                                                     from = "2020-01-01", 
                                                     to = "2022-04-04",
                                                     periodicity = "daily",
                                                     auto.assign=FALSE)} )

closePrices <- lapply(myStocks, Cl)
closePrices <- as.data.frame(do.call(merge, closePrices))
closePrices <- na.omit(log(closePrices/lag(closePrices)))
closePrices <- rownames_to_column(closePrices, "Date")

date <- closePrices["Date"]
date

# will need to probably write a function for this later as at the moment i am just hard coding this

corr_AMZN <- as.data.frame(roll_cor(closePrices$AMZN.Close, closePrices$GSPC.Close, width = 150)) %>% cbind(date) %>% na.omit
corr_NFLX <- as.data.frame(roll_cor(closePrices$NFLX.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_MSFT <- as.data.frame(roll_cor(closePrices$MSFT.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_GOOG <- as.data.frame(roll_cor(closePrices$GOOG.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_NVAX <- as.data.frame(roll_cor(closePrices$NVAX.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_DIS <- as.data.frame(roll_cor(closePrices$DIS.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_WMT <- as.data.frame(roll_cor(closePrices$WMT.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_COST <- as.data.frame(roll_cor(closePrices$COST.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_V <- as.data.frame(roll_cor(closePrices$V.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_FB <- as.data.frame(roll_cor(closePrices$FB.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_AXP <- as.data.frame(roll_cor(closePrices$AXP.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_CMCSA <- as.data.frame(roll_cor(closePrices$CMCSA.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_PFE <- as.data.frame(roll_cor(closePrices$PFE.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit
corr_MRNA <- as.data.frame(roll_cor(closePrices$MRNA.Close, closePrices$GSPC.Close, width = 150))%>% cbind(date) %>% na.omit

corr_AMZN$symbol <- "AMZN"
corr_NFLX$symbol <- "NFLX"
corr_GOOG$symbol <- "GOOG"
corr_NVAX$symbol <- "NVAX"
corr_V$symbol <- "VISA"
corr_FB$symbol <- "FB"
corr_MSFT$symbol <- "MSFT"
corr_PFE$symbol <- "PFE"
corr_MRNA$symbol <- "MRNA"

colnames(corr_AMZN)[1] = "corr"
colnames(corr_NFLX)[1] = "corr"
colnames(corr_MSFT)[1] = "corr"
colnames(corr_GOOG)[1] = "corr"
colnames(corr_NVAX)[1] = "corr"
colnames(corr_DIS)[1] = "corr"
colnames(corr_WMT)[1] = "corr"
colnames(corr_COST)[1] = "corr"
colnames(corr_V)[1] = "corr"
colnames(corr_FB)[1] = "corr"
colnames(corr_AXP)[1] = "corr"
colnames(corr_CMCSA)[1] = "corr"
colnames(corr_PFE)[1] = "corr"
colnames(corr_MRNA)[1] = "corr"

corr_joined = rbind(corr_AMZN, corr_NFLX, corr_GOOG, corr_NVAX, corr_V, corr_FB, corr_MSFT, corr_PFE, corr_MRNA)

corr_joined$Date <- as.Date(corr_joined$Date)
corr_joined

p = ggplot(corr_joined, aes(x=Date, y=corr, colour = symbol)) + geom_line() + labs(x = "Date", y = "rolling correlation (150D period)") + ggtitle("examining correlation of vaccine macufacturers compared to typical stocks") + theme_clean()

ggplotly(p)
```

## Visualization for Price and intertested variables
```{r, warning=FALSE}
df_q2_scatter_plot_Lprice_var_fac = function(plot_df, xvar, xvar_lab, title) {
  p1 = ggplot(data = plot_df, aes(x = xvar, group=location, color = location)) +
    geom_point(aes(y=Lprice)) +
    facet_grid(location ~ .) + 
    ggtitle(title) + xlab(xvar_lab) + 
    ylab("Price") +
    theme_bw()
  p1
}

df_q2$Lprice = log(df_q2$Price/lag(df_q2$Price))
df_q2$Lnew_vaccinations = log(df_q2$new_vaccinations/lag(df_q2$new_vaccinations))
df_q2$Lnew_tests = log(df_q2$new_tests/lag(df_q2$new_tests))
df_q2$Lnew_cases = log(df_q2$new_cases/lag(df_q2$new_cases))

df_q2_scatter_plot_Lprice_var_fac(df_q2, df_q2$Lnew_tests, "new_tests", "log return(Price)~Log return(new_tests)")
df_q2_scatter_plot_Lprice_var_fac(df_q2, df_q2$Lnew_vaccinations, "new_vaccinations", "Log return(Price)~Log return(new_vaccinations)")
df_q2_scatter_plot_Lprice_var_fac(df_q2, df_q2$Lnew_cases, "new_vaccinations", "Log return(Price)~Log return(new_cases)")
```

### Visualization of new vacination, tests, cases over time
```{r}
#colnames(df_q2)
df_q2_date = df_q2 %>% drop_na(new_cases, new_tests, new_vaccinations) %>% group_by(Date) %>% summarize(avg_price = mean(Price), avg_new_cases = mean(new_cases), avg_new_tests = mean(new_tests), avg_new_vac = mean(new_vaccinations))
p1 = ggplot(data = df_q2_date, aes(x = Date)) + geom_line(aes(y=avg_new_cases), color = 'red') +
  geom_line(aes(y=avg_new_vac), color = 'light green') +
  geom_line(aes(y=avg_new_tests), color = 'light blue') +
  ggtitle("covid data plot") + xlab("date") +
  ylab("values")+
  theme_bw()
ggplotly(p1)
```

### Correlation Matrix
```{r}
df_q2_cor = df_q2 %>% drop_na(new_vaccinations, new_tests, new_cases, Price) %>% select(new_vaccinations, new_tests, new_cases, Price)
cor(df_q2_cor)
qtlcharts::iplotCorr(df_q2_cor)
```


## beta function for each country
```{r, eval=FALSE}
#covid_clean
df_q2_beta = df_q2 %>% drop_na(Lprice2, Lnew_cases2)
df_q2_chn = df_q2_beta %>% filter(location == "China")
df_q2_jpn = df_q2_beta %>% filter(location == "Japan")
df_q2_usa = df_q2_beta %>% filter(location == "United States")
df_q2_ind = df_q2_beta %>% filter(location == "India")
chn_beta_price_case = beta(df_q2_chn$Lprice2, df_q2_chn$Lnew_cases2)
jpn_beta_price_case = beta(df_q2_jpn$Lprice2, df_q2_jpn$Lnew_cases2)
usa_beta_price_case = beta(df_q2_usa$Lprice2, df_q2_usa$Lnew_cases2)
ind_beta_price_case = beta(df_q2_ind$Lprice2, df_q2_ind$Lnew_cases2)
aus_beta_price_case = beta(df_q2_aus$Lprice2, df_q2_aus$Lnew_cases2)
plot(jpn_beta_price_case,type="l")
plot(chn_beta_price_case,type="l", col = "red")
plot(aus_beta_price_case,type="l", col = "light blue")
plot(ind_beta_price_case,type="l", col = "light green")
plot(usa_beta_price_case,type="l", col = "yellow")
```


# Question3 

## Building the model
```{r}
glm1= glm(Price ~ new_vaccinations + new_tests + new_cases,data = df_q2_cor)
summary(glm1)
```

## Assumption checking
```{r}
plot(glm1, which = 1:2)
```
```{r}
mean(df_q2_cor$Price)
median(df_q2_cor$Price)
quantile(df_q2_cor$Price, 0.25)
```
```{r}
#df_q2_cor$level <- ifelse(df_q2_cor$Price < 1978.28, "low", "high")

df_q2_cor <- df_q2_cor %>% mutate(level = case_when((df_q2_cor$Price <= 1978.28) ~ "Low",
                                                    ((df_q2_cor$Price > 1978.28) & (df_q2_cor$Price < 6511.446)) ~ "Medium",
                                         
                                                    (df_q2_cor$Price > 6511.446) ~ "High"
))
df_q2_cor
```



```{r}
set.seed(1)
split = 0.8
## 80% 20% train/test
trainIndex <- createDataPartition(df_q2_cor$level, p = split, list = FALSE)
data_train <- df_q2_cor[trainIndex, ]
data_test <- df_q2_cor[-trainIndex, ]
##naive bayes model
model = train(df_q2_cor$level ~ df_q2_cor$new_vaccinations+df_q2_cor$new_tests+df_q2_cor$new_cases+df_q2_cor$Price, data = data_train)
```


```{r}
# define training control
train_control <- trainControl(method="cv", number=10)
# train the model
model <- train(df_q2_cor$level~., data=df_q2_cor, trControl=train_control, method="nb")
```

#Marin

```{r}
table(df_q2_cor$level)
```

```{r}
df_q2_cor$level <- as.factor(df_q2_cor$level)
df_q2_cor
```

```{r,warning=FALSE}
# load the library
library(caret)
# define training control
train_control <- trainControl(method="boot", number=100)
# train the model
model <- train(level~., data=df_q2_cor, trControl=train_control, method="nb")
# summarize results
print(model)
```

```{r,warning=FALSE}
# define training control
train_control <- trainControl(method="repeatedcv", number=10, repeats=3)
# train the model
model <- train(level~., data=df_q2_cor, trControl=train_control, method="nb")
# summarize results
print(model)
```
