---
title: "Group Project 4 COVID"
author: 'covid_p11'
date: '2022-03-28'
output:
  html_document:
    toc: true
    number_sections: true
    theme: flatly
    highlight: tango
    toc_float: true
    toc_depth: 4
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(maps)
library(plotly)
library(caret)
library(dplyr)
library(lubridate)
library("scales")
library("magrittr")
library("tidyr")
library(pkr)
```

Topic: COVID-19 New cases
visualisation dashboard: Shiny app

# NEW FINAL JOIN CODE BY MAXIM

```{r}
##########################
# THIS IS THE FINAL JOIN #
#######################################################################
# use covid_joined for your work and filter it to country if you need #
#######################################################################

covid = read.csv("owid-covid-data.csv")
sp500 = read.csv("SPY Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for US
TOPIX = read.csv("TOPIX Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for Japan
ASX200 = read.csv("S&P_ASX 200 Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for Australia
NSEI = read.csv("Nifty 50 Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for India
SSEC = read.csv("Shanghai Composite Historical Data.csv", fileEncoding = 'UTF-8-BOM') # this is for China

covid2 <- covid

covid2$date = as.character(ymd(covid2$date))
sp500$Date <- as.character(mdy(sp500$Date))
TOPIX$Date <- as.character(mdy(TOPIX$Date))
ASX200$Date <- as.character(mdy(ASX200$Date))
NSEI$Date <- as.character(mdy(NSEI$Date))
SSEC$Date <- as.character(mdy(SSEC$Date))

colnames(covid2)[4] = "Date"

covid_US <- covid2 %>% filter(location == "United States")
covid_AUS <- covid2 %>% filter(location == "Australia")
covid_IND <- covid2 %>% filter(location == "India")
covid_JPN <- covid2 %>% filter(location == "Japan")
covid_CHN <- covid2 %>% filter(location == "China")

df_1 = inner_join(sp500, covid_US, by = "Date")
df_2 = inner_join(TOPIX, covid_JPN, by = "Date")
df_3 = inner_join(ASX200, covid_AUS, by = "Date")
df_4 = inner_join(NSEI, covid_IND, by = "Date")
df_5 = inner_join(SSEC, covid_CHN, by = "Date")

covid_joined <- rbind(df_1, df_2, df_3, df_4, df_5)

covid_joined$Date <- ymd(covid_joined$Date)
covid_joined$Price <- as.numeric(gsub(",","",covid_joined$Price))
covid_joined$Open <- as.numeric(gsub(",","",covid_joined$Open))
covid_joined$High <- as.numeric(gsub(",","",covid_joined$High))
covid_joined$Low <- as.numeric(gsub(",","",covid_joined$Low))

# example analysis for US
US_data = covid_joined[covid_joined$location == "United States",]
plot(US_data$Price ~ US_data$Date)
```


# CLEANING AND (OLD JOIN CODE) (not going to use but will leave it because i think some of the code currently depends on it, will fix in the meeting together)

```{r}
covid = read.csv("owid-covid-data.csv")
ASX200 = read.csv("S&P_ASX 200 Historical Data.csv", fileEncoding = 'UTF-8-BOM') # This is for Australia
```

# data clean

```{r, eval=FALSE}
#Dropping NA
covid_clean = covid %>% drop_na(new_cases, new_cases_smoothed, new_vaccinations, new_vaccinations_smoothed, new_vaccinations_smoothed_per_million, population, population_density, median_age, extreme_poverty, total_vaccinations, hospital_beds_per_thousand, human_development_index, new_deaths, new_tests, weekly_icu_admissions_per_million, weekly_icu_admissions)

#This code changes the negative case values in the data set to zero
covid_clean$new_cases[covid_clean$new_cases < 0] <- 0
```

```{r}
#Dropping NA
covid_AUS <- covid %>% filter(location == "Australia")
covid_clean_AUS = covid_AUS %>% drop_na(new_cases, new_cases_smoothed, new_vaccinations, new_vaccinations_smoothed, new_vaccinations_smoothed_per_million, population, population_density, median_age, extreme_poverty, total_vaccinations, hospital_beds_per_thousand, human_development_index, new_deaths, new_tests, total_tests, total_cases)

#This code changes the negative case values in the data set to zero
covid_clean_AUS$new_cases[covid_clean_AUS$new_cases < 0] <- 0

#glimpse(covid_clean)
#covid_clean$date = as.Date(covid_clean$date)
#max(covid_clean$date)
#unique(covid_clean$location)
```

# the old join

```{r}
covid2 <- covid_clean_AUS

ASX200$date = mdy(ASX200$Date)

covid2$date = ymd(covid2$date)

covid2 <- covid2 %>%
  transform(covid2, date = as.character(date)) %>% 
  select(date, new_cases, new_deaths, location, new_vaccinations, new_tests, population, population_density, total_tests, total_vaccinations, total_cases)

ASX200$date <- as.character(ASX200$date)

ASX200$date <- as.character(ASX200$date)

covid_AUS <- covid2 %>% filter(location == "Australia")

df_3 = inner_join(ASX200, covid_AUS, by = "date")

df_3 = inner_join(ASX200, covid_AUS, by = "date")

df_aus <- df_3

df_aus$date = as.Date(df_aus$date)

df_aus$Price = as.numeric(gsub(",","",df_aus$Price))
```

# Q2 a

```{r}
df_aus = df_aus[order(as.Date(df_aus$date, format="%d/%m/%Y")),]

colnames(df_aus)[7] = 'Change'

#write.csv(df_aus,"df_aus.csv")
```

## Price ~ new_cases

```{r, warning=FALSE}
#glimpse(df_aus)
df_aus_subset = df_aus %>% select(Price, new_cases)

M0 = lm(Price ~ new_cases, data = df_aus_subset) # Null model
summary(M0)
```

## Price ~ new_vaccinations

```{r}
# df_aus_subset = df_aus %>% select(Price, new_deaths) %>% filter(new_deaths != 0)
```

## New deaths and prices 

```{r}
df_aus_subset = df_aus %>% select(Price, new_deaths, new_vaccinations, new_cases)

M1 = lm(Price ~ poly(new_vaccinations, degree=2), data=df_aus_subset)
summary(M1)
```

```{r}
ggplot(data = df_aus_subset, mapping = aes(x = Price, y = new_deaths)) + geom_point() + ggtitle("Prices of stocks and against new death") + ylab("New deaths") + xlab("Price")
```

## Price ~ new_vaccinations

```{r}
ggplot(data = df_aus, mapping = aes(x = new_vaccinations, y = Price)) + geom_point() + ggtitle("Prices of stocks against new vaccinations") + xlab("New vaccinations") + ylab("Price")
```

## Price ~ New tests

```{r}
df_aus_subset = df_aus %>% select(Price, new_tests)
M2 = lm(Price ~ poly(new_tests, degree=2), data=df_aus_subset)
summary(M2)
```

```{r}
ggplot(data = df_aus, mapping = aes(x = new_tests, y = Price)) + geom_point() + ggtitle("Prices of stocks against new tests") + xlab("New tests") + ylab("Price")
```

```{r, eval=FALSE}
ggplot(data = df_aus, mapping = aes(x = Price, y = population)) + geom_point() + ggtitle("Prices of stocks against population") + ylab("New tests") + xlab("Price")
ggplot(data = df_aus, mapping = aes(x = Price, y = population_density)) + geom_point() + ggtitle("Prices of stocks against population density") + ylab("New tests") + xlab("Price")
ggplot(data = df_aus, mapping = aes(x = date, y = population_density)) + geom_line() + ggtitle("Prices of stocks against population density") + ylab("New tests") + xlab("Price") # bad
summary(df_aus$Price)
```

```{r}
df_aus$Price_mul30 = df_aus$Price*30
p1 = ggplot(data = df_aus) + geom_line(aes(x=date, y = new_cases), color = "red") + geom_line(aes(x=date, y = new_vaccinations), color = "light green") + geom_line(aes(x=date, y = new_tests), color = "light blue") +
  geom_line(aes(x=date, y = Price_mul30), color = "blue") + theme_bw()

df_aus$Price_mul30 = df_aus$Price*30
p1 = ggplot(data = df_aus) + geom_line(aes(x=date, y = new_cases), color = "red") + geom_line(aes(x=date, y = new_vaccinations), color = "light green") + geom_line(aes(x=date, y = new_tests), color = "light blue")+ theme_bw()
ggplotly(p1)
```


```{r, eval=FALSE}
#autoplot(M0, which = 1:2)
#autoplot(M1, which = 1:2)
#autoplot(M2, which = 1:2)
```

```{r}
M = lm(Price ~ polym(new_tests, new_vaccinations, degree=2, raw=TRUE), data=df_aus)
summary(M)
ggplot(data = df_aus) + geom_point(aes(x = new_tests, y = Price), color = "light blue") +
  geom_point(aes(x = new_vaccinations, y = Price), color = "light green") +
  ggtitle("Prices of stocks against new tests and new_vaccinations") + ylab("Price") +theme_classic()
```

- The behaviour of a dependent variable is explained by a linear, or curvilinear, additive relationship between the dependent variable and a set of k independent variables (xi, i=1 to k).
- The relationship between the dependent variable and any independent variable is linear or curvilinear.
- The independent variables do no depend on each other too.
- The errors are independent, normally distributed with mean zero and a constant variance.

## Log Price vs grwth rate (give up) 

### Calculate grwth rate

```{r, eval=FALSE}
covid_clean_AUS <- df_aus
percentage_tests <- (covid_clean_AUS$new_tests/ covid_clean_AUS$total_tests)*100
vec_percentage_tests <- c(percentage_tests)
covid_clean_AUS <-cbind(covid_clean_AUS, perc_new_tests = vec_percentage_tests)

covid_clean_AUS <- covid_clean_AUS
percentage_cases <- (covid_clean_AUS$new_cases/ covid_clean_AUS$total_cases)*100
vec_percentage_cases <- c(percentage_cases)
covid_clean_AUS <-cbind(covid_clean_AUS, perc_new_cases = vec_percentage_cases)

covid_clean_AUS <- covid_clean_AUS
percentage_vacc <- (covid_clean_AUS$new_vaccinations/ covid_clean_AUS$total_vaccinations)*100
vec_percentage_vacc <- c(percentage_vacc)
covid_clean_AUS <-cbind(covid_clean_AUS, perc_new_vacc = vec_percentage_vacc)
df_aus_Q2 <- covid_clean_AUS

df_aus_Q2$lprice = log(df_aus_Q2$Price)
```

### Visulization

```{r, eval=FALSE}
ggplot(data = df_aus_Q2) + geom_point(aes(x = perc_new_tests, y = lprice), color = "light blue") +
  geom_point(aes(x = perc_new_vacc, y = lprice), color = "light green") +
  geom_point(aes(x = perc_new_cases, y = lprice), color = "red") +
  ggtitle("Prices of stocks againsts growth rate of new cases, new vaccinations and new tests") + xlab("growth rate") + ylab("log(Price)") +theme_classic()
```

Very bad plot....

```{r, eval=FALSE}
df_aus_Q2 = filter(df_aus_Q2, date >= "2021-03-05")

colors <- c("new vaccinations" = "light green", "new cases" = "red", "new tests" = "orange", "log price" = "black")
p1 = ggplot(data = df_aus_Q2, aes(x=date)) + geom_line(aes(y = perc_new_cases, color = "new cases")) + geom_line(aes(y = perc_new_vacc, color = "new vaccinations")) +
  geom_line(aes(y = lprice, color = "log price")) + 
  geom_line(aes(y = perc_new_tests, color = "new tests")) +
  labs(x = "Date",
         y = "(%)",
         color = "Legend") +
  scale_color_manual(values = colors) + theme_bw()
ggplotly(p1)
```

## Slope

```{r}
class(df_aus$date)
Slope(df_aus$date, df_aus$Price)
```

